Interpretable Machine Learning

### 2018-10-23 poster/poster_2018.Rnw
final proofread

\usepackage{hyperref}
\usetheme{Board}
\usecolorstyle[colorPalette=PurpleGrayBlue]{Germany}

\usebackgroundstyle{Default}
\title{Interpretable Machine Learning}
\author{Christoph Molnar  \url{christoph.molnar@stat.uni-muenchen.de}}
\institute{LMU Munich / Department of Statistics / Prof. Dr. Bernd Bischl}
%\titlegraphic{\includegraphics[scale=0.1]{figure/logo.png}}
\usetitlestyle{Empty}
% \colorlet{titlefgcolor}{black}
% \colorlet{titlebgcolor}{colorOne}

@@

<<ice, echo=FALSE, fig.width=12, fig.height=6>>=
eff = FeatureEffect$new(pred, "size", method = "pdp")
eff.res = eff$results
ggplot(eff.res)  + geom_line(aes(x=size, y = .y.hat), size = 3) +
  scale_x_continuous("Size of living area") + 
  scale_y_continuous("Predicted rent") +
  theme(aspect.ratio=2/4, text = element_text(size=35),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank())
@

\textit{If we vary the size of the living area, we observe how the predicted rente increases with increasing size.}

%Caveat: ICE and PDP have trouble when the inputs are strongly correlated.
%In this case, use Accumulated Local Effects instead \cite{apley2016visualizing}.
}
\block{Feature Importance}{
Feature importance \cite{Fisher2018} tells us how much model error increases when we shuffle its values in the data (= "destroying" the relationship between the feature and the outcome).
The greater the increase in error, the more important the feature is.
<<feature-importance, echo=FALSE, fig.width=12, fig.height=8>>=
imp = FeatureImp$new(pred, "mae")

@@

\textbf{Counterfactual Explanation} \\
A counterfactual explanation of a prediction describes the smallest changes to the feature values (inputs) that change the prediction to a predefined output. \\
\textit{How do the inputs for the 989 Euro apartment have to change so that the predicted rent is over 1100 Euros?
Answer (one of many possible ones): If cats were allowed, the predicted rent would be \Sexpr{pred.new} Euros.}
\\ \\
\textbf{Shapley Values} \\
From a game theory point of view, the feature values are players in a cooperative game who receive the value of the prediction as a payout.
The Shapley Values method \cite{vstrumbelj2014explaining} splits the difference between the prediction and the average prediction fairly among the feature values.

@@

@
\textit{The predicted value of the apartment is 989Euros, the average prediction of all apartments is \Sexpr{avg.pred}. The negative difference is explained by its small size and the ban on cats. The good location has a positive effect on the predicted rent.}

@@

df2 = df
df2$rent = cut(df$rent, breaks = quantile(df$rent, probs = seq(from = 0, to = 1, length.out = 20)))
df2$size = cut(df$size, breaks = quantile(df$size, probs = seq(from = 0, to = 1, length.out = 5)))
rule = JRip(rent ~ ., data = df2)
extract.rules.jrip(rule)
@
\textit{
IF $90m^2\leq \text{size} < 110m^2$  AND location $=$ Good THEN the rent is between 1540 and 1890 EUR}

@@

\textbf{Local Surrogate Models}\cite{ribeiro2016should} \\
To explain an individual prediction, replace a  complex model with a locally weighted, interpretable model (e.g. decision tree).
Data points that are close to the data point of interest get a high weight.

<<lime, echo=FALSE>>=
# locmod = LocalModel$new(pred, k = 2)
# locmod$explain(df[1,1:3])
# plot(locmod)
knitr::include_graphics("figure/local-surrogate.png")
@
}

>>>>

\usepackage{hyperref}
% \definecolorpalette{BlueGrayOrange}{
%     \definecolor{colorOne}{HTML}{C0C0C0}
%     \definecolor{colorTwo}{HTML}{CCCCCC}
%     \definecolor{colorThree}{HTML}{009440}
% }
\definecolorpalette{LMU}{
    \definecolor{colorTwo}{HTML}{CCCCCC}
    \definecolor{colorThree}{HTML}{C0C0C0}
    \definecolor{colorOne}{HTML}{009440}
}

\usetheme{Board}
\usecolorstyle[colorPalette=LMU]{Germany}

% \usebackgroundstyle{Default}

\title{Interpretable Machine Learning}
\author{Christoph Molnar  \url{christoph.molnar@stat.uni-muenchen.de}}
\institute{LMU Munich / Department of Statistics / Prof. Dr. Bernd Bischl}
%\titlegraphic{\includegraphics[scale=0.1]{figure/logo.png}}
\usetitlestyle{Autumn}
\colorlet{titlefgcolor}{black}
\colorlet{titlebgcolor}{white}

@@

<<ice, echo=FALSE, fig.width=12, fig.height=6>>=
eff = FeatureEffect$new(pred, "size", method = "pdp")
eff.res = eff$results
ggplot(eff.res)  + geom_line(aes(x=size, y = .y.hat), size = 3) +
  scale_x_continuous("Size of living area") + 
  scale_y_continuous("Average predicted rent") +
  theme(aspect.ratio=2/4, text = element_text(size=35),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank())
@

\textit{If we vary the size of the living area, we observe how the predicted rent increases with increasing size.}

%Caveat: ICE and PDP have trouble when the inputs are strongly correlated.
%In this case, use Accumulated Local Effects instead \cite{apley2016visualizing}.
}
\block{Feature Importance}{
Feature importance \cite{Fisher2018} tells us how much the model error increases when we shuffle the values of a feature in the data (= "destroying" the relationship between the feature and the outcome).
The greater the increase in error, the more important the feature is.
<<feature-importance, echo=FALSE, fig.width=12, fig.height=8>>=
imp = FeatureImp$new(pred, "mae")

@@

\textbf{Counterfactual Explanation} \\
A counterfactual explanation describes the smallest changes to a data points feature values (inputs)  that change the prediction to a predefined output. \\
\textit{How do the inputs for the 989 Euro apartment have to change so that the predicted rent is over 1100 Euros?
Answer (one of many possible ones): If cats were allowed, the predicted rent would be \Sexpr{pred.new} Euros.}
\\ \\
\textbf{Shapley Values} \\
From a game theory point of view, the feature values are players in a cooperative game who receive the value of the prediction as a payout.
The Shapley Value method \cite{vstrumbelj2014explaining} splits the difference between the prediction and the average prediction fairly among the feature values.

@@

@
\textit{The predicted value of the apartment is 989Euros, the average prediction of all apartments is \Sexpr{avg.pred}. The negative difference is explained by the apartments small size and the ban on cats. The good location has a positive effect on the predicted rent.}

@@

df2 = df
df2$rent = cut(df$rent, breaks = quantile(df$rent, probs = seq(from = 0, to = 1, length.out = 20)))
df2$size = cut(df$size, breaks = quantile(df$size, probs = seq(from = 0, to = 1, length.out = 5)))
rule = JRip(rent ~ ., data = df2)
extract.rules.jrip(rule)
@
\textit{
IF $90m^2\leq \text{size} < 110m^2$  AND location $=$ "good" THEN rent is between 1540 and 1890 EUR}

@@

\textbf{Local Surrogate Models}\cite{ribeiro2016should} \\
To explain an individual prediction, replace a  complex model with a locally weighted, interpretable model (e.g. decision tree).
Data points that are close to the data point of interest get a high weight.
\vspace{0.5cm}
<<lime, echo=FALSE>>=
# locmod = LocalModel$new(pred, k = 2)
# locmod$explain(df[1,1:3])
# plot(locmod)
knitr::include_graphics("figure/local-surrogate.png")
@
}
